\documentclass[12pt]{article}
\usepackage{pmmeta}
\pmcanonicalname{VirtuallyCyclicGroup}
\pmcreated{2013-03-22 15:47:15}
\pmmodified{2013-03-22 15:47:15}
\pmowner{yark}{2760}
\pmmodifier{yark}{2760}
\pmtitle{virtually cyclic group}
\pmrecord{11}{37746}
\pmprivacy{1}
\pmauthor{yark}{2760}
\pmtype{Definition}
\pmcomment{trigger rebuild}
\pmclassification{msc}{20F19}
\pmclassification{msc}{20E22}
\pmsynonym{cyclic-by-finite group}{VirtuallyCyclicGroup}
\pmrelated{VirtuallyAbelian}
\pmdefines{virtually cyclic}
\pmdefines{cyclic-by-finite}
\pmdefines{finite-by-cyclic group}
\pmdefines{finite-by-cyclic}

\usepackage{amsfonts}
\usepackage{amsthm}

\newtheorem*{thm*}{Theorem}
\newtheorem*{cor*}{Corollary}

\def\Z{\mathbb{Z}}
\begin{document}
\PMlinkescapeword{corollary}
\PMlinkescapeword{index}
\PMlinkescapeword{infinite}
\PMlinkescapeword{theorem}
\PMlinkescapeword{type}
\PMlinkescapeword{types}

A \emph{virtually cyclic group} is 
a group that has a cyclic subgroup of finite \PMlinkname{index}{Coset}.
Every virtually cyclic group in fact 
has a normal cyclic subgroup of finite index
(namely, the core of any cyclic subgroup of finite index),
and virtually cyclic groups are therefore also known as
\emph{cyclic-by-finite groups}.

A finite-by-cyclic group 
(that is, a group $G$ with a finite normal subgroup $N$
such that $G/N$ is cyclic)
is always virtually cyclic.
To see this, note that a finite-by-cyclic group is either finite,
in which case it is certainly virtually cyclic,
or it is finite-by-$\Z$, 
in which case the \PMlinkname{extension}{GroupExtension}
\PMlinkname{splits}{SemidirectProductOfGroups}.

Finite-by-\PMlinkname{dihedral}{DihedralGroup} groups 
are also virtually cyclic.
In fact, we have the following classification theorem:\cite{hempel}\cite{rdiat}

\begin{thm*}
Groups of the following three types are all virtually cyclic.
Moreover, every virtually cyclic group is of exactly one of these three types.
\begin{itemize}
\item finite
\item finite-by-(infinite cyclic)
\item finite-by-(infinite dihedral)
\end{itemize}
\end{thm*}

As an immediate corollary we have the following result:\cite{macpherson}

\begin{cor*}
Every torsion-free virtually cyclic group is either trivial or infinite cyclic.
\end{cor*}

\begin{thebibliography}{9}
\bibitem{hempel}
 Lemma 11.4 (pages 102--103) in:
 John Hempel,
 {\it 3-Manifolds},
 American Mathematical Society, 2004,
 ISBN 0821836951.
\bibitem{rdiat}
 Page 137 of:
 Alejandro Adem, Jesus Gonzalez, Guillermo Pastor (eds.),
 Recent developments in algebraic topology --- A conference to celebrate
   Sam Gitler's 70th birthday, San Miguel de Allende, Mexico, December 3--6, 2003.
\bibitem{macpherson}
 Lemma 3.2 (pages 225--226) of:
 Dugald Macpherson,
 {\it Permutation Groups Whose Subgroups Have Just Finitely Many Orbits}
 (pages 221--230 in: W.\ Charles Holland (ed.)
  {\it Ordered Groups and Infinite Permutation Groups},
  Kluwer Academic Publishers, 1996).
\end{thebibliography}
%%%%%
%%%%%
\end{document}
